---
title: "Project 1"
author: "Ciara Whipp"
date: "6/10/2020"
output:
  rmarkdown::github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE, message=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DT)
```

## Introduction to JSON Data  

### What is JSON and what is its use?  

JavaScript Object Notation or [JSON](https://www.tutorialspoint.com/json/json_overview.htm#:~:text=JSON%20or%20JavaScript%20Object%20Notation,stands%20for%20JavaScript%20Object%20Notation.) is a text-based, data-interchangable format used to store and transport data. It is often used to transmit data between a server and web applications, and web services and APIs use JSON to provide data. We will see JSON provided by an API in this document.  

### Why use JSON?  

JSON is language independent, so major programming languages are cable of handling JSON data - they can get data out of JSON and turn data into JSON. The intuitive nature of the syntax is easy to read and write, and since the syntax is derived from and similar to existing programming languages, it is easy for machines to generate and parse JSON data. This [video](https://www.youtube.com/watch?v=0IoG-mSvWSo) provides a brief overview of JSON data: what it is, advantages to JSON, and an introduction to JSON syntax.  

### R Packages to Accomodate JSON Data  

There are three major R packages used to deal with JSON data: `jsonlite`, `rjson`, and `RJSONIO`.  
Some helpful functions included in all three packages are `fromJSON()` to convert JSON data into an R object, `toJSON()` to convert an R object into JSON. Addtionally, `flatten` is a useful argument that creates a single non-nested data frame from a nested data frame (common in JSON data retrieved from the web). I will choose to use `jsonlite` since the R documentation for this package is more robust than the doucmentation for the other two packages.  

## Retrieving Data from an API  

The following code creates the base url that can be used to read in data from the National Hockey League (NHL) API.  

```{r baseURL, echo=TRUE, eval=TRUE}
baseURL <- "https://records.nhl.com/site/api/"
```

The following function can be used to access the **franchise** data and **franchise-team-totals** data. Note the `jsonlite` and `httr` packages are required to utilize this function.  

```{r overall, echo=TRUE, eval=TRUE, message=FALSE}
library(jsonlite)
library(httr)
overallData <- function(x){
  fullURL <- paste0(baseURL,x)
  GET(fullURL) %>% 
    content("text") %>% 
    fromJSON(flatten=TRUE) %>% 
    as_tibble()
}
```

Use the function above to return tibbles for the **franchise** data and **franchise-team-totals** data.  

```{r overallData, echo=TRUE, eval=TRUE, message=FALSE}
overallData(x="franchise")
overallData(x="franchise-team-totals")
```

The following function can be used to access **franchise-season-records**, **franchise-goalie-records**, and **franchise-skater-records** for a specified franchise ID.  

```{r franchise, echo=TRUE, eval=TRUE, message=FALSE}
franchiseData <- function(x,ID){
  fullURL <- paste0(baseURL,x,"?cayenneExp=franchiseId=",ID)
  GET(fullURL) %>% 
    content("text") %>% 
    fromJSON(flatten=TRUE) %>% 
    as_tibble()
}
```  

Use the function above to return tibbles for the Capitals franchise (ID=24) for **franchise-season-records**, **franchise-goalie-records**, and **franchise-skater-records**.  

```{r franchiseData, echo=TRUE, eval=TRUE, message=FALSE}
franchiseData(x="franchise-season-records", ID="24")
franchiseData(x="franchise-goalie-records", ID="24")
franchiseData(x="franchise-skater-records", ID="24")
```

## Creating New Variables  

Create a new variable for the **franchise-skater-records** data for the Washington Capitals called goalsPerGame that gives the average number of goals per game by dividing `goals` by `gamesPlayed`. Return a table for the top 6 goals per game statistic.  

```{r goalsTable, echo=TRUE, eval=TRUE, message=FALSE}
skaterCaps <- franchiseData(x="franchise-skater-records", ID="24")$data %>% 
  rename("Last Name"=lastName, "Position Code"=positionCode, "Number of Seasons Played"=seasons) %>%
  mutate("Goals per Game"=goals/gamesPlayed) %>% 
  select(`Last Name`, `Position Code`, `Goals per Game`, `Number of Seasons Played`) %>% 
  arrange(desc(`Goals per Game`))
knitr::kable(head(skaterCaps), digits=2)
```

## Numeric Summaries  

### Contingency Tables  

Continuing with the **franchise-skater-records** data for the Washington Capitals, create a contingency table that give the counts for each position code based on active status.  

```{r contigencyTable1, echo=TRUE, eval=TRUE, message=FALSE}
skaterCaps <- franchiseData(x="franchise-skater-records", ID="24")$data
knitr::kable(table(skaterCaps$activePlayer, skaterCaps$positionCode), 
             caption="Active Status and Position Code Information for the Capitals")
```

Using the **franchise-goalie-records** data for the Washington Capitals, change the values of `mostGoalsAgainstOneGame` to "Good/Fair Performance" for 4 or less goals, "Poor Performance" for 5-8 goals, and "Terrible Performance" for more than 8 goals to represent Goalie Performance, and change the values of `wins` to "Winning Goalie" for goalies with greater than or equal to 175 wins and to "Losing Goalie" for goalies with less than 175 wins to represent Winnning Status. The contingency tables show Goalie Performance and Winning status for non-active and active Capitals' goalies.  

```{r contingencyTable2, echo=TRUE, eval=TRUE, message=FALSE}
goalieData<-franchiseData(x="franchise-goalie-records", ID="24")$data
x<-goalieData$mostGoalsAgainstOneGame
goalieData$mostGoalsAgainstOneGame <- ifelse(x<=4, "Good/Fair Performance",ifelse(x>8, "Terrible Performance", "Poor Performance"))
y<-goalieData$wins
goalieData$wins <- ifelse(y<175, "Losing Goalie","Winning Goalie")
notActive <- goalieData %>% 
  filter(activePlayer == FALSE)
knitr::kable(table(notActive$mostGoalsAgainstOneGame, notActive$wins), 
             caption="Goalie Performance and Winning Status for Non-Active Goalies")
active <- goalieData %>% 
  filter(activePlayer ==TRUE)
knitr::kable(table(active$mostGoalsAgainstOneGame, active$wins), 
             caption="Goalie Performance and Winning Status for Active Goalies")
```

### Summary Statistics Tables  

The tables below give summary statisitics for non-active and active Capitals' goalies (minimum, 1st quantile, median, mean, 3rd quantile, and maximum) for `mostSavesOneGame`, `mostShotsAgainstOneGame`, `mostWinsOneSeason`, and `seasons`.  

```{r summaryStats, echo=TRUE, eval=TRUE, message=FALSE}
notActive <- notActive %>%
  select(mostSavesOneGame, mostShotsAgainstOneGame,
         mostWinsOneSeason, seasons) %>%
  rename("Most Saves in 1 Game"=mostSavesOneGame, 
         "Most Shots Against in 1 Game"=mostShotsAgainstOneGame, 
         "Most Wins in 1 Season"=mostWinsOneSeason,
         "Seasons"=seasons)
knitr::kable(apply(notActive,2,summary), digits=1,
             caption="Summary of Non-Active Capitals' Goalies")
active <- active %>%
    select(mostSavesOneGame, mostShotsAgainstOneGame,
         mostWinsOneSeason, seasons) %>%
    rename("Most Saves in 1 Game"=mostSavesOneGame, 
         "Most Shots Against in 1 Game"=mostShotsAgainstOneGame, 
         "Most Wins in 1 Season"=mostWinsOneSeason,
         "Seasons"=seasons)
knitr::kable(apply(active,2,summary), 
             caption="Summary of Active Capitals' Goalies")
```

## Visuals  

Using the `goalieData` from the tables above, create various types of plots to visualize the data.  

### Bar Plots  

#### One Variable Bar Plots  

The bar plots below represent the counts of Non-active and Active Goalies, of Goalie Perforamance, and Winning Status, respectively.  

```{r barPlots1, echo=TRUE, eval=TRUE, message=FALSE}
g <- ggplot(goalieData)
g + geom_bar(aes(x=activePlayer)) + 
  labs(x="Player Active")
g + geom_bar(aes(x=mostGoalsAgainstOneGame)) + 
  labs(x="Goalie Performance")
g + geom_bar(aes(x=wins)) +
  labs(x="Winning Status")
```

#### Two Variable Side-by-Side Bar Plots  

The first bar plot is a visual representation of the number of Active and Non-Active Goalies for each Goal Performance category. The second bar plot is a visual respresentation fo the number of active and non-active goalies for Winning Goalies and Losing Goalies.

```{r barPlots2, echo=TRUE, eval=TRUE, message=FALSE}
g + geom_bar(aes(x=mostGoalsAgainstOneGame, fill=as.factor(activePlayer)),
             position="dodge") +
  labs(x="Goalie Performance") +
  scale_fill_discrete(name="Active Player")
g + geom_bar(aes(x=wins, fill=as.factor(activePlayer)),
             position="dodge") +
  labs(x="Winning Status") +
  scale_fill_discrete(name="Active Player")
```

#### Three Variable Side-by-Side Bar Plot  

The following bar plots give a visual representation of the number of Winning Goalies and Losing Goalies for Active and Non-Active Goalies for each Goalie Performance category.

```{r barPlots3, echo=TRUE, eval=TRUE, message=FALSE}
g + geom_bar(aes(x=activePlayer, fill=as.factor(wins)), position="dodge") +
  labs(x="Active Player") +
  scale_fill_discrete("Winning Status") +
  facet_wrap(~ mostGoalsAgainstOneGame)
```

### Box Plots  

The following box plot gives visual representation and comparison of the summary statistics for Most Saves in One Game for Active and Non-Active Goalies.  

```{r, boxPlots1, echo=TRUE, eval=TRUE, message=FALSE}
g <- ggplot(goalieData, aes(x=activePlayer, y=mostSavesOneGame))
g + geom_boxplot() +
  geom_point(aes(x=activePlayer, y=mostSavesOneGame,
                 color=activePlayer),position="jitter") +
  scale_color_discrete(name="Active Player") +
  labs(x="Active Player", y="Most Saves in 1 Game")
```

The following box plot gives visual representation and comparison of the summary statistics for Most Wins in One Season for Winning Goalies and Losing Goalies.  

```{r boxPlots2, echo=TRUE, eval=TRUE, message=FALSE}
g <- ggplot(goalieData, aes(x=wins, y=mostWinsOneSeason))
g + geom_boxplot() +
  geom_point(aes(x=wins, y=mostWinsOneSeason,
                 color=wins),position="jitter") +
  scale_color_discrete(name="Winning Status") +
  labs(x="Winning Status", y="Most Wins in 1 Season")
```

### Scatter Plots  

The following scatter shows the relationship between Number of Seasons Played and Most Wins in 1 Season for both Winning Goalies and Losing Goalies.  

```{r scatterPlots1, echo=TRUE, eval=TRUE, message=FALSE}
g <- ggplot(goalieData, aes(x=seasons, y=mostWinsOneSeason))
g + geom_point(aes(color=wins)) +
  scale_color_discrete(name="Winning Status") +
  labs(x="Number of Seasons Played", y="Most Wins in 1 Season")
```

The following scatter plot shows the relationship between Most Shots Against in 1 Game and Most Saves in 1 Game for Active and Non-Active Goalies.  

```{r scatterPlots2, echo=TRUE, eval=TRUE, message=FALSE}
g <- ggplot(goalieData, aes(x=mostShotsAgainstOneGame, y=mostSavesOneGame))
g + geom_point(aes(color=activePlayer)) +
  geom_smooth(aes(group=activePlayer), color="green", method=lm, lwd=.5) +
  scale_color_discrete(name="Active Player") +
  labs(x="Most Shots Against in 1 Game", y="Most Saves in One Game")
```