---
title: "Project 1"
author: "Ciara Whipp"
date: "6/10/2020"
output: 
  rmarkdown::github_document:
    toc: true
---

```{r setup, include=FALSE, message=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(httr)
library(jsonlite)
```

## Introduction to JSON Data  

### What is JSON and what is its use?  

JavaScript Object Notation or [JSON](https://www.tutorialspoint.com/json/json_overview.htm#:~:text=JSON%20or%20JavaScript%20Object%20Notation,stands%20for%20JavaScript%20Object%20Notation.) is a text-based, data-interchangable format used to store and transport data. It is often used to transmit data between a server and web applications, and web services and APIs use JSON to provide data. We will explore JSON provided by an API in this document.  

### Why use JSON?  

JSON is language independent, so major programming languages are cable of handling JSON data - they can get data out of JSON and turn data into JSON. The intuitive nature of the syntax is easy to read and write, and since the syntax is derived from and similar to existing programming languages, it is easy for machines to generate and parse JSON data. This [video](https://www.youtube.com/watch?v=0IoG-mSvWSo) provides a brief overview of JSON data: what it is, advantages to JSON, and an introduction to JSON syntax.  

### R Packages to Accomodate JSON Data  

There are three major R packages used to deal with JSON data: `jsonlite`, `rjson`, and `RJSONIO`.  
Some helpful functions included in all three packages are `fromJSON()` to convert JSON data into an R object, `toJSON()` to convert an R object into JSON. Addtionally, `flatten` is a useful argument that creates a single non-nested data frame from a nested data frame (common in JSON data retrieved from the web). I will choose to use `jsonlite` since the R documentation for this package is more robust than the doucmentation for the other two packages.  

## Retrieving Data from an API  

The following code creates the base url that can be used to read in data from the National Hockey League (NHL) API.  

```{r baseURL, echo=TRUE, eval=TRUE}
baseURL <- "https://records.nhl.com/site/api/"
```

The following function can be used to access **franchise** data and **franchise-team-totals** data.

```{r overall, echo=TRUE, eval=TRUE, message=FALSE}
overallData <- function(x){
  fullURL <- paste0(baseURL,x)
  GET(fullURL) %>% content("text") %>% fromJSON() %>% as_tibble()
}
```

The following function can be used to access **franchise-season-records**, **franchise-goalie-records**, and **franchise-skater-records** for a specified franchise ID. 

```{r franchise, echo=TRUE, eval=TRUE, message=FALSE}
franchiseData <- function(x,ID){
  fullURL <- paste0(baseURL,x,"?cayenneExp=franchiseId=",ID)
  GET(fullURL) %>% content("text") %>% fromJSON() %>% as_tibble()
}
```